---
title: "ST3131"
output: html_document
---
```{r echo=FALSE}
library(tidyverse)
library(readxl)
data <- read.csv("FEV.csv")
data <- select(data,c("Age","FEV","Hgt","Sex","Smoke"))
```
As the ID and Hgt_m are redundant in the dataset, we remove them.
```{r echo=FALSE}
summary(data$FEV)
```

For the variables, Age, Hgt are continuous variables, while Sex and Smoke are indicator variables.
```{r echo=FALSE}
ggplot(data,aes(Age,FEV))+geom_point()+geom_smooth(method="lm")
ggplot(data,aes(Hgt,FEV))+geom_point()+geom_smooth(method="lm")
```
From the plots we can See that there is positive correlation between Age and FEV as well as between Hgt and FEV.

To examine the effect of smoking on FEV, we divide the dataset into smokers and non-smokers.
```{r echo=FALSE}
smoker <- filter(data,Smoke==1)
```

As the youngest smoker is 9 years old, we need to subset the non-smoker dataset by Age>=9 years old.
```{r echo=FALSE}
nonsmoker <- filter(data,Smoke==0,Age>=9)
ggplot(smoker,aes(Age,FEV))+geom_point()+geom_smooth(method="lm")+geom_smooth(data=nonsmoker,method="lm",color="red")
```
From the plot we can see that the best-fit line of non-smoker is steeper, which means that the coefficient is larger.
This means that as Age increases, FEV of non-smokers increases more than that of smokers, which shows that there is
a negative correlation between smoking and FEV.

To examine the effect of different gender on FEV, we divide the dataset into male and female.
```{r echo=FALSE}
female <- filter(data,Sex==0)
male <- filter(data,Sex==1)
ggplot(male,aes(Age,FEV))+geom_point()+geom_smooth(method="lm")+geom_smooth(data=female,method="lm",color="red")
```
From the plot we can see that the best fit line of male is steeper,which means that the coefficient is lager.
This means that as Age increases, FEV of male increases more than that of female.
```{r echo=FALSE}
ggplot(data,aes(FEV))+geom_histogram()
```
The response could be a normal distribution.
```{r echo=FALSE}
library(corrplot)
M <- cor(data[,c(1,3,4,5)])
corrplot(M, method = "color", diag=F,type="upper",
         title = "Variable Correlation Plot",
         tl.col='#4477AA',tl.cex=0.7,
         mar = c(2, 1, 3, 1))
```
Age and Hgt, Age and Smoke has relatively strong relationship

```{r echo=FALSE}
model<- lm(FEV ~ Hgt + Age + Sex + Smoke,data)
summary(model)
anova(model)
plot(model$fitted.values,rstandard(model))
abline(h=0)
```
From the summary table, we can see that Hgt, Age, Sex are significant variables as the p-value is very small. Smoke is not a significant variable. This may not be the case as shown by the earlier plot and by common sense, smoking would affect FEV.
We can see that the residual is not randomly distributed around 0 which contradicts the assumption of a linear model. Therefore we need a better model. 
```{r echo=FALSE}
model_2 <- lm(log(FEV) ~ Hgt+Age+Sex+Smoke,data)
summary(model_2)
anova(model_2)
plot(model_2$fitted.values,rstandard(model_2))
abline(h=0)
```

```{r echo=FALSE}
model_inter <- lm(FEV~1, data = data)
fwd <- step(model_inter, direction = c("forward"), scope = FEV ~ Hgt+Age+Sex+Smoke)
summary(fwd)
bwd <- step(model_2, direction = c("backward"))
summary(bwd)
```
From the result of variable selection, we can keep all 4 variables.

From the summary table of the second model which is the improved model. We can see that the p-value of every variable is significant while Hgt and Age are more significant compared to Sex and Smoke.
The residual is randomly distributed around 0 therefore it is a good estimation of linear model.

Therefore the final model I selected would be model_2 which is a multiple linear regression model between ln(FEV) and Hgt, Age, Sex and Smoke. We can see that 1 inch increase in height would increase ln(FEV) by 0.042783, 1 year increase in age would increase ln(FEV) by 0.023387. Male has a higher ln(FEV) of 0.029236 compared to Female and non-smoker has a higher ln(FEV) of 0.046015 compared to smokers. The result reflects the relationships of different variables with FEV shown in Part I.